Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-28 paper.tex

Various science gateways have been developed, including frameworks
used as building blocks to assemble science gateways. The main
frameworks are Apache Airavata~\cite{marru2011apache}, the Catania
Science Gateway Framework~\cite{Ardizzone2012} and
WS-PGRADE/gUSE~\cite{Kacsuk2012}. Numerous science gateway instances
were built using such
frameworks~\cite{kacsuk2014science,ardizzone2012decide}. Other gateway
instances were also built independently from such specialized framework, using
regular web development tools and libraries to access distributed
systems. CBRAIN and VIP are two such science gateways, from which we
gathered experience about the architectures described in this paper.

CBRAIN\footnote{\url{http://cbrain.mcgill.ca}}~\cite{SHER-14} is an
open-source\footnote{\url{https://github.com/aces/cbrain}}, web-based collaborative platform addressing the
challenges of data-heavy, computation-intensive neuroimaging
research. It offers transparent web-based access to remote data
sources, distributed computing sites, and an array of processing and
visualization tools, within a secure environment. CBRAIN is deployed
at a wide range of computing sites, including a network of national HPC
centers in Canada, a cluster in Germany, and on numerous local research
servers. CBRAIN is a task-centric platform where various workflow engines
can be deployed on the infrastructure as any other task would. As February 2016,
CBRAIN has served over 420 users across 50 cities in 22 countries and has a
national computing allocation of 660 CPU years per annum.

The Virtual Imaging Platform
(VIP\footnote{\url{http://www.creatis.insa-lyon.fr/vip}})~\cite{GLAT-13}
is a web portal for medical simulation and image data analysis. It
leverages computing and storage resources available in the biomed
Virtual Organization of the European Grid Infrastructure (EGI) to
offer an open service to academic researchers worldwide. VIP
completely masks the infrastructure to enable a transparent user
experience. It uses the MOTEUR workflow engine~\cite{GLAT-13} and the
DIRAC pilot-job framework~\cite{tsaregorodtsev2008dirac} to execute
applications on EGI. Almost 900 users were registered in VIP by
April 2016 and the average CPU consumption between January 2013 and
March 2016 was 35 CPU years per month.

The architectures described hereafter also build on experience with
the SHIWA Simulation Platform~\cite{terstyanszky2014enabling}. The
SHIWA Simulation Platform is a WS-PGRADE/gUSE science gateway that was
developed in the European projects
SHIWA\footnote{\url{http://www.shiwa-workflow.eu}} and
ER-flow\footnote{\url{http://www.erflow.eu}}. It focuses on workflow
interoperability for various disciplines, in particular through the
building of meta-workflows, i.e. workflows that are built out of other
workflows. It implements the so-called Coarse-Grained Interoperability
framework where workflow engines are integrated in the science gateway
as other applications and invoked through a dedicated service. The
Fine-Grained Interoperability framework was also developed in the
SHIWA project, where conversions between workflow languages are
achieved through a common intermediate
representation~\cite{plankensteiner-prodan-etal:2013}.

>>>>

Various science gateways have been developed, including frameworks
used as building blocks to assemble science gateways. The main
frameworks are Apache Airavata~\cite{marru2011apache}, the Catania
Science Gateway Framework~\cite{Ardizzone2012} and
WS-PGRADE/gUSE~\cite{Kacsuk2012}. Numerous science gateway instances
were built using such
frameworks~\cite{kacsuk2014science,ardizzone2012decide}. Other gateway
instances were also built independently from such specialized framework, using
regular web development tools and libraries to access distributed
systems, for instance CBRAIN~\cite{SHER-14} and VIP~\cite{GLAT-13}.
